---
title: "Data simulations: pleiotropy + individual-level correlations"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
library(knitr)
opts_chunk$set(dev = "png", dpi = 192, fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

# Include

```{r inc}
library(tidyverse)
library(magrittr)

library(Matrix)
library(MASS)

library(nlme)
library(lme4)

library(ggplot2)
library(cowplot)
theme_set(theme_cowplot(font_size = 10))

# change to your path here, e.g. based on `nodemame`
path <- switch(Sys.info()[['nodename']],
  "~/git/hemostat/polycor/", 'gen-sflaptop' = '/media/sf9/ExtraDrive1/Dropbox/Research_170228/covQTL/bin/polycor/')
  
src <- file.path(path, "relmer.R")
source(src)

src <- file.path(path, "simFunctions.R")
source(src)
```

# Parameters

```{r par}
testing <- FALSE

h2_cov <- 0.3
h2_add <- 0.3
vals_h2_t1 <- c(1e-4, 0.1, 0.2, 0.3)
rhog <- 0.5

nrep <- 10
vals_n <- 5 * c(100, 500, 1e3) / nrep
vals_sim <- 1:5

if(testing) {
  nrep <- 2
  vals_n <- c(10, 50) / nrep
  vals_sim <- 1:2
}
```

# Functions

```{r run_models}
run_models <- function(models, dat, G)
{
  lapply(models, function(model) {
    tab <- switch(model,
      "bivar" = {
        bdat <- gather(dat, tname, tvalue, trait1, trait2)
          mod <- relmer(tvalue ~ -1 + tname + (0 + tname | id) + (0 + tname | rid), 
            data = bdat, relmat = list(id = G), calc.derivs = FALSE)
          
          # cov <- VarProp(mod) %>% as.data.frame %>% 
          #   filter(grp == "id" & var1 == "tnametrait1" & var2 == "tnametrait2") %$% prop
          vf <- VarProp(mod) %>% as.data.frame %>% filter(grp == "id")
          cov <- filter(vf, grp == "id" & var1 == "tnametrait1" & var2 == "tnametrait2") %$% vcov
          rhog <- cov / sqrt(vf$vcov[1]*vf$vcov[2])

          add <- VarProp(mod) %>% as.data.frame %>% 
            filter(grp == "id" & var1 == "tnametrait2" & is.na(var2)) %$% prop
                    
          # data_frame(cov = cov, add = add)
          data_frame(rhog = rhog, add = add)
      },
      "slope_add" = {
          mod <- relmer(trait2 ~ (1|rid) + (0 + trait1|id), data = dat, 
            relmat = list(id = G, rid = G))
        
          cov <- VarProp(mod) %>% as.data.frame %>% filter(grp == "id") %$% prop
          add <- VarProp(mod) %>% as.data.frame %>% filter(grp == "rid") %$% prop
          
          data_frame(cov = cov, add = add)
      },
      stop("error in switch: unknown `model`"))
        
    mutate(tab, model = model)
  }) %>% bind_rows
}
```

# Simulation 1: h2.cov + h2.add

| Data Simulation | Bi-variate Model | Random-slope Model |
|--|--|--|
| Random-slope (h2.cov = `r h2_cov`; h2.add = `r h2_add`) | v | v |

```{r sim1, cache = T}
grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim) %>%
  as_tibble

tab <- lapply(seq(nrow(grid)), function(i)
{
  n <- grid$n[i]
  sim <- grid$sim[i]
  
  message(" -", i, "/", nrow(grid), "\n")
  
  simdat <- genBivarCov.simulatePairKin(n_ind = n, n_perInd = nrep, 
    h2.cov = h2_cov, h2.add = h2_add,
    h2.add.t1 = 1e-4, rhog = 0,
    postproc = TRUE)

  dat <- simdat$pheno %>% as_tibble
  G <- simdat$G
  
  models <- c("bivar", "slope_add")
  run_models(models, dat, G) %>%
    mutate(
      true_cov = h2_cov, true_add = h2_add,
      n = n, nrep = nrep)
}) %>% bind_rows %>%
  mutate(N = factor(n, levels = unique(n), labels = unique(paste0("N = ", n, " x ", nrep))))
```  

```{r pltot_sim1, echo = F, cache = T}
ptab <- mutate(tab, 
    bivar = ifelse(model == "bivar", cov, NA), 
    cov = ifelse(model == "bivar", NA, cov)) %>%
  gather(comp, var, add, cov, bivar)
  
ggplot(ptab, aes(y = var, color = model)) + geom_boxplot() + facet_grid(comp ~ N) + geom_hline(yintercept = c(0, h2_cov, h2_add), linetype = 3) + theme(legend.position = "top") + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) 
```

# Simulation 2: h2.cov + h2.add + h2.add.t1

| Data Simulation | Bi-variate Model | Random-slope Model |
|--|--|--|
| Random-slope (h2.cov = `r h2_cov`; h2.add = `r h2_add`; h2.add.t1) | v | v |

```{r sim2, cache = T}
grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim,
  h2.add.t1 = vals_h2_t1) %>%
  as_tibble

tab <- lapply(seq(nrow(grid)), function(i)
{
  n <- grid$n[i]
  sim <- grid$sim[i]
  h2_add_t1 <- grid$h2.add.t1[i]
  
  message(" -", i, "/", nrow(grid), "\n")
  
  simdat <- genBivarCov.simulatePairKin(n_ind = n, n_perInd = nrep, 
    h2.cov = h2_cov, h2.add = h2_add,
    h2.add.t1 = h2_add_t1, rhog = 0,
    postproc = TRUE)

  dat <- simdat$pheno %>% as_tibble
  G <- simdat$G
  
  models <- c("bivar", "slope_add")
  run_models(models, dat, G) %>%
    mutate(
      true_cov = h2_cov, true_add = h2_add, true_h2_add_t1 = h2_add_t1,
      n = n, nrep = nrep)
}) %>% bind_rows %>%
  mutate(N = factor(n, levels = unique(n), labels = unique(paste0("N = ", n, " x ", nrep))))
```  

```{r pltot_sim2, echo = F, cache = T}
ptab <- mutate(tab, 
    bivar = ifelse(model == "bivar", cov, NA), 
    cov = ifelse(model == "bivar", NA, cov)) %>%
  gather(comp, var, add, cov, bivar)

ggplot(ptab, aes(x = as.factor(true_h2_add_t1), y = var, color = model)) + geom_boxplot() + facet_grid(comp ~ N) + geom_hline(yintercept = c(0, h2_cov, h2_add), linetype = 3) + theme(legend.position = "top") 
```

# Simulation 3: h2.cov + h2.add + h2.add.t1 + rhog

| Data Simulation | Bi-variate Model | Random-slope Model |
|--|--|--|
| Random-slope (h2.cov = `r h2_cov`; h2.add = `r h2_add`; h2.add.t1; rhog = `r rhog`) | v | upward bias (when #repeats is low) |

NOTE: the upward bias is noticable when the number of repeats is 4. On the figure below, the number of repeats is 10, and the bias is small. 

TODO: We might try a different data-simulation model, where `cor(t1, t2)` is afected by only the genetic additive part of `t1` rather than additive+bi-variate.

Data simulation in Simulation 3:

```
trait1 ~ g.add.t1 + g.bivar + e.t1
trait2 ~ g.cov * trait1 + g.add.t2 + g.bivar + e.t2
```

An updated data simulation model:

```
trait1 ~ g.add.t1 + g.bivar + e.t1
trait2 ~ g.cov * [g.add.t1 + e.t1] + g.add.t2 + g.bivar + e.t2
```

```{r sim3, cache = T}
grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim,
  h2.add.t1 = vals_h2_t1) %>%
  as_tibble

tab <- lapply(seq(nrow(grid)), function(i)
{
  n <- grid$n[i]
  sim <- grid$sim[i]
  h2_add_t1 <- grid$h2.add.t1[i]
  
  message(" -", i, "/", nrow(grid), "\n")
  
  simdat <- genBivarCov.simulatePairKin(n_ind = n, n_perInd = nrep, 
    h2.cov = h2_cov, h2.add = h2_add,
    h2.add.t1 = h2_add_t1, rhog = rhog,
    postproc = TRUE)

  dat <- simdat$pheno %>% as_tibble
  G <- simdat$G
  
  models <- c("bivar", "slope_add")
  run_models(models, dat, G) %>%
    mutate(
      true_cov = 1e-4, true_add = h2_add, true_h2_add_t1 = h2_add_t1,
      n = n, nrep = nrep)
}) %>% bind_rows %>%
  mutate(N = factor(n, levels = unique(n), labels = unique(paste0("N = ", n, " x ", nrep))))
```  

```{r pltot_sim3, echo = F, cache = T}
ptab <- mutate(tab, 
    bivar = ifelse(model == "bivar", rhog, NA), 
    cov = ifelse(model == "bivar", NA, cov)) %>%
  gather(comp, var, add, cov, bivar)


ggplot(ptab, aes(x = as.factor(true_h2_add_t1), y = var, color = model)) + geom_boxplot() + facet_grid(comp ~ N, scales = 'free_y') + geom_hline(yintercept = c(0, h2_cov, h2_add, rhog), linetype = 3) + theme(legend.position = "top") 
```

# Simulation 4: bi-variate h2.add + h2.add.t1 + rhog

| Data Simulation | Bi-variate Model | Random-slope Model |
|--|--|--|
| Random-slope (h2.add = `r h2_add`; h2.add.t1; rhog = 0.9) | v | v |

TODO: to check what kind of variance in estimated by uni- and bi-variate model (under bi-variate model simulation). It seems that

- bi-variate model (including random-slope model) picks up `g.add.t2` 
- uni-variate model picks up `g.add.t2 + g.bivar`

```
trait1 ~ g.add.t1 + g.bivar + e.t1
trait2 ~ g.cov * trait1 + g.add.t2 + g.bivar + e.t2
```

```{r sim4, cache = T}
grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim,
  h2.add.t1 = vals_h2_t1) %>%
  as_tibble

tab <- lapply(seq(nrow(grid)), function(i)
{
  n <- grid$n[i]
  sim <- grid$sim[i]
  h2_add_t1 <- grid$h2.add.t1[i]
  
  message(" -", i, "/", nrow(grid), "\n")
  
  simdat <- genBivarCov.simulatePairKin(n_ind = n, n_perInd = nrep, 
    h2.cov = 1e-4, h2.add = h2_add,
    h2.add.t1 = h2_add_t1, rhog = 0.9,
    postproc = TRUE)

  dat <- simdat$pheno %>% as_tibble
  G <- simdat$G
  
  models <- c("bivar", "slope_add")
  run_models(models, dat, G) %>%
    mutate(
      true_cov = h2_cov, true_add = h2_add, true_h2_add_t1 = h2_add_t1,
      n = n, nrep = nrep)
}) %>% bind_rows %>%
  mutate(N = factor(n, levels = unique(n), labels = unique(paste0("N = ", n, " x ", nrep))))
```  

```{r pltot_sim4, echo = F, cache = T}
ptab <- mutate(tab, 
    bivar = ifelse(model == "bivar", rhog, NA), 
    cov = ifelse(model == "bivar", NA, cov)) %>%
  gather(comp, var, add, cov, bivar)

ggplot(ptab, aes(x = as.factor(true_h2_add_t1), y = var, color = model)) + geom_boxplot() + facet_grid(comp ~ N, scales = 'free_y') + geom_hline(yintercept = c(0, h2_add, 0.9), linetype = 3) + theme(legend.position = "top") 
```
