---
title: "Data simulations: pleiotropy + individual-level correlations. Genotype based. No repeated observations"
author: "Simon Forsberg"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
    toc: true
    keep_md: true
---


```{r options, echo = F}
library(knitr)
opts_chunk$set(dev = "png", dpi = 192, fig.path = "../figures/genoSim_n1perInd/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

Simulations based on actual genotypes in outbred but stratified populations. Same thing as previous genotype based simulations (polycor/issues/13-post-ASHG/R/genotypeSimulations_currentLme4qtl.Rmd) but with no repeated observations per individual. Instead, these simulations include more individuals.

## Include

```{r inc}
library(tidyverse)
library(magrittr)

library(Matrix)
library(MASS)

library(nlme)
library(lme4)

library(cowplot)
library(wesanderson)
theme_set(theme_cowplot(font_size = 10))

# change to your path here, e.g. based on `nodemame`
path <- switch(Sys.info()[['nodename']],
               "~/git/hemostat/polycor/", 'gen-sflaptop' = '/media/sf9/ExtraDrive1/Dropbox/Research_170228/covQTL/bin/polycor/',
               'gen-comp1.princeton.edu' = '/Genomics/ayroleslab2/simon/covProj/bin/polycor',
               'gen-comp2.princeton.edu' = '/Genomics/ayroleslab2/simon/covProj/bin/polycor')

src <- file.path(path, "modelFitting.R")
source(src)

src <- file.path(path, "simFunctions.R")
source(src)

src <- file.path(path, "plotFunctions.R")
source(src)
```

## Parameters
These parameters are common to all the simulations

```{r par}
testing <- T

n_loci <- 200
nrPops <- 10
Fst <- .3
models <- c("bivar", "slope_add_12")

#Sample sizes and number of simulations
nrep <- 1
vals_n <- c(5000, 10000)
vals_sim <- 1:3

if(testing) {
  nrep <- 1
  vals_n <- c(20, 40)
  vals_sim <- 1:2
}
```


# Simulation 1: h2.cov + h2.add.t2

```{r sim1, cache = T}
#Parameters
h2_cov <- c(0.3, 0.6)
h2_add_t1 <- 0
h2_add_t2 <- c(1e-4, 0.1, 0.2, 0.3)
rhog <- 0
resFile <- '../results/genoSim_n1perInd/sim1.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog)
  SimRes <- data.frame(grid, runtime_slope = NA, runtime_bivar = NA,
                       h2.add.t2_slopeMod = NA, 
                       h2.add.t1_bivar = NA, h2.add.t2_bivar = NA,
                       h2.cov_slopeMod = NA, rhog_bivar = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.add.t2 <- grid$h2_add_t2[i]
    h2.cov <- grid$h2_cov[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2_add_t1, h2.add.t2 = h2.add.t2,
                                               h2.cov = h2.cov, rhog = rhog, 
                                               nrPops = nrPops, Fst = Fst, postproc = T)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G)

    #Store results
    SimRes$h2.add.t2_slopeMod[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slope_add_12']
    SimRes$h2.add.t1_bivar[i] <- simdat.modFit$h2.add.t1[simdat.modFit$model == 'bivar']
    SimRes$h2.add.t2_bivar[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'bivar']
    SimRes$h2.cov_slopeMod[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slope_add_12']
    SimRes$rhog_bivar[i] <- simdat.modFit$rhog[simdat.modFit$model == 'bivar']
    SimRes$runtime_slope[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slope_add_12']
    SimRes$runtime_bivar[i] <- simdat.modFit$runtime[simdat.modFit$model == 'bivar']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  


## h2.add.t2
* x-axis = h2.add.t2 true
* y-axis = h2.add.t2 estimated

```{r plot_sim1_addT2, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)
```

## h2.cov & rhog
* x-axis = h2.add.t2 true
* y-axis = h2.cov (slope_add model), rhog (bivar model)

Dashed lines indicate simulated values for h2.cov and rhog
```{r plot_sim1_h2cov_rhog, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('slope_add    n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))
```



# Simulation 2: h2.cov + h2.add.t1 + h2.add.t2

```{r sim2, cache = T}
#Parameters
h2_cov <- c(0.3, 0.6)
h2_add_t1 <- .3
h2_add_t2 <- c(1e-4, 0.1, 0.2, 0.3)
rhog <- 0
resFile <- '../results/genoSim_n1perInd/sim2.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog)
  SimRes <- data.frame(grid, runtime_slope = NA, runtime_bivar = NA,
                       h2.add.t2_slopeMod = NA, 
                       h2.add.t1_bivar = NA, h2.add.t2_bivar = NA,
                       h2.cov_slopeMod = NA, rhog_bivar = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.add.t2 <- grid$h2_add_t2[i]
    h2.cov <- grid$h2_cov[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2_add_t1, h2.add.t2 = h2.add.t2,
                                               h2.cov = h2.cov, rhog = rhog, 
                                               nrPops = nrPops, Fst = Fst, postproc = T)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G)

    #Store results
    SimRes$h2.add.t2_slopeMod[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slope_add_12']
    SimRes$h2.add.t1_bivar[i] <- simdat.modFit$h2.add.t1[simdat.modFit$model == 'bivar']
    SimRes$h2.add.t2_bivar[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'bivar']
    SimRes$h2.cov_slopeMod[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slope_add_12']
    SimRes$rhog_bivar[i] <- simdat.modFit$rhog[simdat.modFit$model == 'bivar']
    SimRes$runtime_slope[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slope_add_12']
    SimRes$runtime_bivar[i] <- simdat.modFit$runtime[simdat.modFit$model == 'bivar']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  

## h2.add.t2
* x-axis = h2.add.t2 true
* y-axis = h2.add.t2 estimated

```{r plot_sim2_addT2, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)
```

## h2.cov & rhog
* x-axis = h2.add.t2 true
* y-axis = h2.cov (slope_add model), rhog (bivar model)

Dashed lines indicate simulated values for h2.cov and rhog

```{r plot_sim2_h2cov_rhog, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('slope_add    n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))
```

# Simulation 3: h2.cov + h2.add.t1 + h2.add.t2 + rhog

```{r sim3, cache = T}
#Parameters
h2_cov <- c(0.3, 0.6)
h2_add_t1 <- .3
h2_add_t2 <- c(1e-4, 0.1, 0.2, 0.3)
rhog <- .5
resFile <- '../results/genoSim_n1perInd/sim3.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog)
  SimRes <- data.frame(grid, runtime_slope = NA, runtime_bivar = NA,
                       h2.add.t2_slopeMod = NA, 
                       h2.add.t1_bivar = NA, h2.add.t2_bivar = NA,
                       h2.cov_slopeMod = NA, rhog_bivar = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.add.t2 <- grid$h2_add_t2[i]
    h2.cov <- grid$h2_cov[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2_add_t1, h2.add.t2 = h2.add.t2,
                                               h2.cov = h2.cov, rhog = rhog, 
                                               nrPops = nrPops, Fst = Fst, postproc = T)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G)

    #Store results
    SimRes$h2.add.t2_slopeMod[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slope_add_12']
    SimRes$h2.add.t1_bivar[i] <- simdat.modFit$h2.add.t1[simdat.modFit$model == 'bivar']
    SimRes$h2.add.t2_bivar[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'bivar']
    SimRes$h2.cov_slopeMod[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slope_add_12']
    SimRes$rhog_bivar[i] <- simdat.modFit$rhog[simdat.modFit$model == 'bivar']
    SimRes$runtime_slope[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slope_add_12']
    SimRes$runtime_bivar[i] <- simdat.modFit$runtime[simdat.modFit$model == 'bivar']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  

## h2.add.t2
* x-axis = h2.add.t2 true
* y-axis = h2.add.t2 estimated

```{r plot_sim3_addT2, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.4), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)
```

## h2.cov & rhog
* x-axis = h2.add.t2 true
* y-axis = h2.cov (slope_add model), rhog (bivar model)

Dashed lines indicate simulated values for h2.cov and rhog

```{r plot_sim3_h2cov_rhog, echo = F, cache = T}
cols <- wes_palette('Darjeeling2', 2)

par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('slope_add    n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.9), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))
```

# Simulation 4: h2.cov + h2.add.t1 + h2.add.t2 + rhog

Same as simulation 3 but with rhog = 0.9

```{r sim4, cache = T}
#Parameters
h2_cov <- c(0.3, 0.6)
h2_add_t1 <- .3
h2_add_t2 <- c(1e-4, 0.1, 0.2, 0.3)
rhog <- .9
resFile <- '../results/genoSim_n1perInd/sim4.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog)
  SimRes <- data.frame(grid, runtime_slope = NA, runtime_bivar = NA,
                       h2.add.t2_slopeMod = NA, 
                       h2.add.t1_bivar = NA, h2.add.t2_bivar = NA,
                       h2.cov_slopeMod = NA, rhog_bivar = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.add.t2 <- grid$h2_add_t2[i]
    h2.cov <- grid$h2_cov[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2_add_t1, h2.add.t2 = h2.add.t2,
                                               h2.cov = h2.cov, rhog = rhog, 
                                               nrPops = nrPops, Fst = Fst, postproc = T)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G)

    #Store results
    SimRes$h2.add.t2_slopeMod[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slope_add_12']
    SimRes$h2.add.t1_bivar[i] <- simdat.modFit$h2.add.t1[simdat.modFit$model == 'bivar']
    SimRes$h2.add.t2_bivar[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'bivar']
    SimRes$h2.cov_slopeMod[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slope_add_12']
    SimRes$rhog_bivar[i] <- simdat.modFit$rhog[simdat.modFit$model == 'bivar']
    SimRes$runtime_slope[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slope_add_12']
    SimRes$runtime_bivar[i] <- simdat.modFit$runtime[simdat.modFit$model == 'bivar']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  

## h2.add.t2
* x-axis = h2.add.t2 true
* y-axis = h2.add.t2 estimated

```{r plot_sim4_addT2, echo = F, cache = T}
par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,.45), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.45), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.45), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.add.t2_slopeMod', 'h2.add.t2_bivar')], ylab = '', xlab = '', legend.cex = 3, spacer2 = 2, ylim = c(0,.45), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = unique(SimRes$h2_add_t2), lty = 3)
```

## h2.cov & rhog
* x-axis = h2.add.t2 true
* y-axis = h2.cov (slope_add model), rhog (bivar model)

Dashed lines indicate simulated values for h2.cov and rhog

```{r plot_sim4_h2cov_rhog, echo = F, cache = T}
cols <- wes_palette('Darjeeling2', 2)

par(mfrow = c(2,2), mar = c(3,2,2,1))
incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.text = c('slope_add', 'bivar'), legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,1), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[1] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,1), cex.axis = 1, cex.lab = 1)
title(paste('slope_add    n =', vals_n[1], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[1]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,1), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[1]))
abline(h = c(h2_cov[1], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))

incl <- SimRes$n == vals_n[2] & SimRes$h2_cov == h2_cov[2]
plot.contBoxes(x = SimRes$h2_add_t2[incl], y = SimRes[incl, c('h2.cov_slopeMod', 'rhog_bivar')], legend.title = NULL, ylab = '', xlab = '', spacer2 = 2, ylim = c(0,1), cex.axis = 1, cex.lab = 1)
title(paste('n =', vals_n[2], 'x', nrep), line = 1.2)
mtext(side = 3, paste('h2.cov =', h2_cov[2]))
abline(h = c(h2_cov[2], rhog), lty = 2, lwd = 2, col = wes_palette('Darjeeling2', 2))
```

# Runtime
```{r plot_runtime, echo = F, cache = T}
resFiles <- list.files('../results/genoSim_n1perInd/', full.names = T)
SimRes.all <- data.frame()
for (i in 1:length(resFiles)) {
  load(resFiles[i])
  SimRes.all <- rbind(SimRes.all, SimRes)
}

par(mfrow = c(2,1), mar = c(3,4,2,2))
boxplot(SimRes.all$runtime_slope[SimRes.all$n == vals_n[1]], 
        SimRes.all$runtime_bivar[SimRes.all$n == vals_n[1]], 
        main = paste('n =', vals_n[1], 'x', nrep), xlab = '', ylab = 'Minutes', las = 2)
axis(side = 1, at = 1:2, labels = c('slope_add', 'bivar'))

boxplot(SimRes.all$runtime_slope[SimRes.all$n == vals_n[2]], 
        SimRes.all$runtime_bivar[SimRes.all$n == vals_n[2]], 
        main = paste('n =', vals_n[2], 'x', nrep), xlab = '', ylab = 'Minutes', las = 2)
axis(side = 1, at = 1:2, labels = c('slope_add', 'bivar'))
```

Average runtimes (minutes):

| slope_add **n = `r vals_n[1]`**x10 | bivar **n = `r vals_n[1]`**x10 | slope_add **n = `r vals_n[2]`**x10 | bivar **n = `r vals_n[2]`**x10 |
|--------------|--------------|--------------|--------------|
| `r round(mean(SimRes.all$runtime_slope[SimRes.all$n == vals_n[1]]), digits = 3)` | `r round(mean(SimRes.all$runtime_bivar[SimRes.all$n == vals_n[1]]), digits = 3)`| `r round(mean(SimRes.all$runtime_slope[SimRes.all$n == vals_n[2]]), digits = 3)` | `r round(mean(SimRes.all$runtime_bivar[SimRes.all$n == vals_n[2]]), digits = 3)` |


