---
title: "Learning basics about correlations"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(dev = "png", dpi = 192, fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

```{r inc, cache = F, echo = F}
library(pander)
```

```{r settings, cache = F, echo = F}
panderOptions('table.style', 'rmarkdown')

panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)

panderOptions('table.alignment.default', 'left')
```

# Include custom code

```{r inc_custom, cache = F}
library(Matrix)
library(MASS)
library(lme4)

library(tidyverse)
library(magrittr)

library(ggplot2)
library(cowplot)
theme_set(theme_cowplot())

path <- switch(Sys.info()[['nodename']],
  "~/git/hemostat/polycor/" )
  
file.path(path, "relmer.R") %>% source
file.path(path, "simFunctions.R") %>% source
```

# Two pleitropic traits

We quickly run a single round of data simulation and conclusions differ from simulation to simulation.
Multiple runs and larger smaple sizes are needed to evaluate the accurcy of estimates.

## Two pleitropic traits + bi-variate model

Given the two traits are simulated under a bi-variate model with a substantial portion of genetic covariance,
what estimates of heritabilities are expected from uni-variate models?

Simulate two pleitropic traits:
 
```{r simbivar, cache = T}
set.seed(1)
simdat <- pleiotropy.simulatePairKin(n_ind = 500, n_perInd = 1, 
  h2 = 0.5, overlap = 0.9)

dat <- simdat$pheno %>% as_tibble %>%
  mutate(
    trait1sc = as.numeric(scale(trait1)), trait2sc = as.numeric(scale(trait2)),
    id = as.factor(as.character(id)), rid = id)
bdat <- gather(dat, tname, tvalue, trait1sc, trait2sc)

G <- simdat$G
```

Fit the standard bi-variate model (weights):


```{r bivar, cache = T, dependson = "simbivar"}
mod <- relmer(tvalue ~ -1 + tname + (0 + tname|id) + (0 + tname|rid), 
  bdat, relmat = list(id = G), 
  weights = rep(1e10, nrow(bdat)), calc.derivs = FALSE)
mod
```


Fit the standard bi-variate model (no weights/redundant residual random effects):


```{r bivar_noweights, cache = T, dependson = "simbivar"}
mod <- relmer(tvalue ~ -1 + tname + (0 + tname|id) + (0 + tname|rid), 
  bdat, relmat = list(id = G), 
  calc.derivs = FALSE)
mod
```

Fit uni-variate models for each trait and estimates of heritability:

```{r univar, cache = T, dependson = "simbivar"}
mod1 <- relmer(trait1 ~ (1|id), dat, relmat = list(id = G)) 
mod1 %>% VarProp %>% .[1, "prop"]

mod2 <- relmer(trait2 ~ (1|id), dat, relmat = list(id = G)) 
mod2 %>% VarProp %>% .[1, "prop"]
```


## Two random-sloped traits + bi-variate model

We simulate two traits under a extreme scenario for bi-variate model
(random-slope data simulation model):

- no additive variance for any trait;
- random-slope genetic covariance.

What estimates of heritability are produced by uni-variate models?

Simulate two traits:

```{r simcov, cache = T}
set.seed(1)
simdat <- genCov.simulatePairKin(n_ind = 500, n_perInd = 1, 
  h2.cov = 0.9, h2.add = 0)

dat <- simdat$pheno %>% as_tibble %>%
  mutate(
    trait1sc = as.numeric(scale(trait1)), trait2sc = as.numeric(scale(trait2)),
    id = as.factor(as.character(id)), rid = id)
bdat <- gather(dat, tname, tvalue, trait1sc, trait2sc)

G <- simdat$G
```

Fit the standard bi-variate model:

- in some runs: the overall genetic component of variance is close to zero;
- in other runs: the trait2-specific heritabilty seems different from zero.

```{r bivar2, cache = T, dependson = "simcov"}
mod <- relmer(tvalue ~ -1 + tname + (0 + tname|id) + (0 + tname|rid), 
  bdat, relmat = list(id = G), 
  calc.derivs = FALSE)
mod
```

Fit uni-variate models for each trait and estimates of heritability:

- the estimates of heritability for each trait is zero.

```{r univar2, cache = T, dependson = "simcov"}
mod1 <- relmer(trait1 ~ (1|id), dat, relmat = list(id = G)) 
mod1 %>% VarProp %>% .[1, "prop"]

mod2 <- relmer(trait2 ~ (1|id), dat, relmat = list(id = G)) 
mod2 %>% VarProp %>% .[1, "prop"]
```

# Population vs. within individual correlation

## Two pleitropic traits + repeats 

```{r bivar_rep_cor, cache = T}
set.seed(1)
simdat <- pleiotropy.simulatePairKin(n_ind = 10, n_perInd = 100, 
  h2 = 0.95, overlap = 0.95)

dat <- simdat$pheno %>% as_tibble %>%
  mutate(
    trait1sc = as.numeric(scale(trait1)), trait2sc = as.numeric(scale(trait2)))

ggplot(dat, aes(trait1sc, trait2sc)) + geom_point() + geom_smooth(method = "lm", color = "red") + geom_hline(yintercept = 0, linetype = 3) + geom_abline(linetype = 3) + labs(title = "Population-level corr.")

ggplot(filter(dat, id %in% 1:4), aes(trait1sc, trait2sc)) + geom_point() + geom_smooth(method = "lm", color = "blue") + geom_hline(yintercept = 0, linetype = 3) + geom_abline(linetype = 3) + facet_wrap(~ id)  + labs(title = "Individual-level corr.")
```

## Two random-slope traits + repeats 

```{r slope_rep_cor, cache = T}
set.seed(1)
simdat <- genCov.simulatePairKin(n_ind = 10, n_perInd = 100, 
  h2.cov = 0.95, h2.add = 0)

dat <- simdat$pheno %>% as_tibble %>%
  mutate(
    trait1sc = as.numeric(scale(trait1)), trait2sc = as.numeric(scale(trait2)))

ggplot(dat, aes(trait1sc, trait2sc)) + geom_point() + geom_smooth(method = "lm", color = "red") + geom_hline(yintercept = 0, linetype = 3) + geom_abline(linetype = 3) + labs(title = "Population-level corr.")

ggplot(filter(dat, id %in% 1:4), aes(trait1sc, trait2sc)) + geom_point() + geom_smooth(method = "lm", color = "blue") + geom_hline(yintercept = 0, linetype = 3) + geom_abline(linetype = 3) + facet_wrap(~ id) + labs(title = "Individual-level corr.")
```

```{r ind_slopes, dependson = -1}
simdat$pheno$g.cov %>% unique %>% head(4) %>% round(2)
```

