---
title: 'Data simulations: binary traits'
author: "Simon Forsberg"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: null
  theme: default
subtitle: GxEMM & GE random slope model
keep_md: yes
urlcolor: blue
---

```{r options, echo = F}
library(knitr)
opts_chunk$set(dev = "png", dpi = 192, fig.path = "../figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

## Generative model in simulations

$y_1 = G*g_{add1} + e_1$

It is then discretized by simply dividing the range of $y_1$ into two pieces of equal length, coded as 0,1 (`as.numeric(cut(y1, breaks = 2)) - 1`)

$y_2 = G*g_{add2} + (G*g_{cov} + e_{cov})*y_1 + e_2$

**Where:**

* G is the genotype matrix
* $g_{add1} \sim N(0, h^2_{add1}/n_{loci})$ is the per SNP additive effect on y1
* $g_{add2} \sim N(0, h^2_{add2}/n_{loci})$ is the per SNP additive effect on y2
* $g_{cov} \sim N(0, h^2_{cov}/n_{loci})$ is the per SNP effect on the slope
* $e_{cov} \sim N(0, \sigma^2_{cov})$ is the per individual environmental effect on the slope
* $e_1 \sim N(0, 1 - h^2_{add1})$
* $e_2 \sim N(0, 1 - h^2_{add2} - h^2_{cov} - \sigma^2_{cov})$
* $r_g$ describes the fraction of g_add1 and g_add2 that are the same, eg pleiotropy. ($r_g = 0$ here)

## Estimated Models

#### GE random slope model
Both environmental and genetic random slopes:

$y_2 = \mu + b_{g,int} + (\beta + b_{g,slope} + b_{e,slope})*y_1 + e$

**Where:**

* $\mu$ and $\beta$ are fixed effects
* $b_{g,int} \sim N(0,K\sigma^2_{g,int})$
* $b_{g,slope} \sim N(0,K\sigma^2_{g,slope})$
* $b_{e,slope} \sim N(0,I\sigma^2_{e,slope})$
* $e \sim N(0,I\sigma^2_{e})$
* K is the kinship matrix, estimated from G

#### GxEMM

* gtype=iid
* etype=iid

Details [here](https://www.sciencedirect.com/science/article/abs/pii/S0002929719304628)



```{r inc, echo=F}
library(tidyverse)
library(magrittr)

library(Matrix)
library(MASS)

library(nlme)
library(lme4)

library(cowplot)
library(wesanderson)
library(ggpubr)
library(GxEMM)

theme_set(theme_cowplot(font_size = 10))

# change to your path here, e.g. based on `nodemame`
path <- switch(Sys.info()[['nodename']],
               "~/git/hemostat/polycor/", 'gen-sflaptop' = '/media/sf9/ExtraDrive1/Dropbox/Research_170228/covQTL/bin/polycor/',
               'gen-comp1.princeton.edu' = '/Genomics/ayroleslab2/simon/covProj/bin/polycor',
               'gen-comp2.princeton.edu' = '/Genomics/ayroleslab2/simon/covProj/bin/polycor')

src <- file.path(path, "modelFitting.R")
source(src)

src <- file.path(path, "simFunctions.R")
source(src)

src <- file.path(path, "plotFunctions.R")
source(src)

ldak_loc  <- file.path(path, "../gxemm/ldak5.linux")
```


## Parameters

```{r par}
testing <- F

n_loci <- 400 #number of loci
nrPops <- 10  #number of subpopulations
Fst <- .3     #degree of stratification
models <- c("slopeGandE_add_12", "gxemm_giid_eiid")

#Sample sizes and number of simulations
nrep <- 1       #n observations
vals_n <- 1000  #N Individuals
vals_sim <- 1:5 #Simulations per parameter combination

h2_cov <- c(0, 0.2, 0.4)
h2_add_t1 <- 0
h2_add_t2 <- 0
sigma_cov <- c(0, 0.2, 0.4)
rhog <- 0

if(testing) {
  nrep <- 1
  vals_n <- 100
  vals_sim <- 1:2
}
```


# Simulation 1: $h^2_{cov} + \sigma^2_{cov}$

No repeated observations

```{r sim1, cache = T}
resFile <- '../results/sim_gxemm_slopGE_binaryY1.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog, sigma_cov = sigma_cov)
  SimRes <- data.frame(grid, runtime_slopeGE = NA, runtime_eiid_gxemm = NA, 
                       h2.add.t2_slopeMod_GE = NA, 
                       h2.cov_slopeMod_GE = NA, 
                       sigma.cov_slopeMod_GE = NA,
                       gxemm_eiid_h2.hom = NA,
                       gxemm_eiid_h2.het = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.cov <- grid$h2_cov[i]
    h2.add1 <- grid$h2_add_t1[i]
    h2.add2 <- grid$h2_add_t2[i]
    sigma.cov <- grid$sigma_cov[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2.add1, h2.add.t2 = h2.add2,
                                               h2.cov = h2.cov, rhog = rhog, e.cov = sigma.cov,
                                               nrPops = nrPops, Fst = Fst, postproc = T, trait1.binary = T)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G, ldak_loc = ldak_loc)
    
    #Store results
    SimRes$runtime_slopeGE[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$runtime_eiid_gxemm[i] <- simdat.modFit$runtime[simdat.modFit$model == 'gxemm_giid_eiid']
    SimRes$h2.add.t2_slopeMod_GE[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$h2.cov_slopeMod_GE[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$sigma.cov_slopeMod_GE[i] <- simdat.modFit$e.cov_12[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$gxemm_eiid_h2.hom[i] <- simdat.modFit$h2.hom_gxemm[simdat.modFit$model == 'gxemm_giid_eiid']
    SimRes$gxemm_eiid_h2.het[i] <- simdat.modFit$h2.het_gxemm[simdat.modFit$model == 'gxemm_giid_eiid']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  

## Results

* x-axes = simulated $h^2_{cov}$ 
* y-axes = estimates of $h^2_{cov}$, $\sigma^2_{cov}$ (from GE random slope model), and $h^2_{het}$ (from GxEMM) 

```{r plot_h2cov_ecov_gxemm, echo = F, cache = T}
cols1 <- grep('cov_', colnames(SimRes))
cols2 <- grep('gxemm.*h2.het', colnames(SimRes))
cols <- c(cols1, cols2)

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[1]

SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_1 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[1]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.2
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[2]
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_2 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[2]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.4
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[3]
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_3 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[3]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')


legend <- cowplot::get_legend(g_1)
ggarrange(g_1 + theme(legend.position = "none"), 
          g_2 + theme(legend.position = "none"), 
          g_3 + theme(legend.position = "none"),
          legend)
```


## Runtime
```{r plot_runtime_slopeGE_gxemm, echo = F, cache = T}
boxplot(SimRes$runtime_slopeGE, SimRes$runtime_eiid_gxemm, xaxt = 'n', ylab = '')
axis(side = 1, at = 1:2, labels = c('Random Slope GE', 'GxEMM eiid'), cex.axis = 1.5)
mtext('Minutes', side = 2, cex = 2, line = 2)
```





# Simulation 2: $h^2_{cov} + \sigma^2_{cov}$

3 observations per individual observations

```{r sim2, cache = T}
models <- "slopeGandE_add_12"
nrep <- 3       #n observations
y1Binary <- c(T,F)

resFile <- '../results/sim2_nrep3_slopGE_binaryY1.RData'

if(!file.exists(resFile)){
  grid <- expand.grid(n = as.integer(vals_n), sim = vals_sim, h2_add_t1 = h2_add_t1, h2_add_t2 = h2_add_t2, h2_cov = h2_cov, rhog = rhog, sigma_cov = sigma_cov, y1Binary = y1Binary)
  SimRes <- data.frame(grid, runtime_slopeGE = NA, 
                       h2.add.t2_slopeMod_GE = NA, 
                       h2.cov_slopeMod_GE = NA, 
                       sigma.cov_slopeMod_GE = NA)
  
  for(i in 1:nrow(grid)){
    n <- grid$n[i]
    sim <- grid$sim[i]
    h2.cov <- grid$h2_cov[i]
    h2.add1 <- grid$h2_add_t1[i]
    h2.add2 <- grid$h2_add_t2[i]
    sigma.cov <- grid$sigma_cov[i]
    y1Binary <- grid$y1Binary[i]
    message(i, "/", nrow(grid), "\n")
    
    #Run simulation
    simdat <- genBivarCov.simulatePair_outbred(n_ind = n, n_perInd = nrep, 
                                               h2.add.t1 = h2.add1, h2.add.t2 = h2.add2,
                                               h2.cov = h2.cov, rhog = rhog, e.cov = sigma.cov,
                                               nrPops = nrPops, Fst = Fst, postproc = T, trait1.binary = y1Binary)
    
    #Fit models
    simdat.modFit <- run_models(models, simdat$pheno, simdat$G, ldak_loc = ldak_loc)
    
    #Store results
    SimRes$runtime_slopeGE[i] <- simdat.modFit$runtime[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$h2.add.t2_slopeMod_GE[i] <- simdat.modFit$h2.add.t2[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$h2.cov_slopeMod_GE[i] <- simdat.modFit$h2.cov_12[simdat.modFit$model == 'slopeGandE_add_12']
    SimRes$sigma.cov_slopeMod_GE[i] <- simdat.modFit$e.cov_12[simdat.modFit$model == 'slopeGandE_add_12']
  }
  save(list = 'SimRes', file = resFile)
}else
  load(resFile)
```  


## Results

* x-axes = simulated $h^2_{cov}$ 
* y-axes = estimates of $h^2_{cov}$, $\sigma^2_{cov}$ 

### Binary y1

```{r plot_h2cov_ecov_nrep3_binary, echo = F, cache = T}
cols <- grep('cov_', colnames(SimRes))

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[1] & SimRes$y1Binary

SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_1 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[1]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.2
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[2] & SimRes$y1Binary
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_2 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[2]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.4
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[3] & SimRes$y1Binary
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_3 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[3]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')


legend <- cowplot::get_legend(g_1)
ggarrange(g_1 + theme(legend.position = "none"), 
          g_2 + theme(legend.position = "none"), 
          g_3 + theme(legend.position = "none"),
          legend)
```

### Continuous y1

```{r plot_h2cov_ecov_nrep3_cont, echo = F, cache = T}
cols <- grep('cov_', colnames(SimRes))

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[1] & !SimRes$y1Binary

SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_1 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[1]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.2
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[2] & !SimRes$y1Binary
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_2 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[2]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')

#h2_add_t1 = 0, h2_add_t2 = 0, sigma_cov = 0.4
incl <- SimRes$h2_add_t1 == h2_add_t1[1] & SimRes$h2_add_t2 == h2_add_t2[1] & SimRes$sigma_cov == sigma_cov[3] & !SimRes$y1Binary
SimRes.long <- pivot_longer(SimRes[incl,], cols = cols)
g_3 <- ggplot(data = SimRes.long, aes(y = value, x = factor(h2_cov), color = name)) + 
  geom_boxplot() +
  theme(legend.title=element_text(size=30, face = 'bold'), 
        legend.text=element_text(size=10), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        plot.title = element_text(hjust = 0.5, size = 10),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) + 
  ylab('Estimate') + xlab(expression('Simulated' ~ h[cov]^2)) +
  guides(colour=guide_legend(title="", override.aes = list(size = 2)) ) +
  ggtitle(bquote(sigma[e.cov]^2 ~ '=' ~ .(sigma_cov[3]) ~ h[add1]^2 ~ '=' ~ .(h2_add_t1[1]) ~ h[add2]^2 ~ '=' ~ .(h2_add_t2[1]))) +
  geom_hline(yintercept = unique(SimRes$h2_cov), linetype = 'dashed')


legend <- cowplot::get_legend(g_1)
ggarrange(g_1 + theme(legend.position = "none"), 
          g_2 + theme(legend.position = "none"), 
          g_3 + theme(legend.position = "none"),
          legend)
```


## Runtime
```{r plot_runtime_slopeGE_nrep3, echo = F, cache = T}
boxplot(SimRes$runtime_slopeGE, xaxt = 'n', ylab = '')
axis(side = 1, at = 1, labels = 'Random Slope GE', cex.axis = 1.5)
mtext('Minutes', side = 2, cex = 2, line = 2)
```


